---
title: "Preprocess the raw behavioral data"
output: html_notebook
---

This script aimed to process the shuffled blocks of experiment 2 and 3. The experiment 1 has no shuffled blocks. And the script is aimed to process the raw data.



This script has three parts:

- clean the data of experiment 1

- select the data of the task 2 from experiment 2&3

- select the data of the task 3 from experiment 2&3



load package

```{r}
# load package
# if 'pacman' is not exist, this package can be installed
if (!require('pacman')) {install.packages('pacman')}
# the package 'pacman' can use 'p_load' to load existed package or install package
library(pacman)
pacman::p_load('tidyverse', # this package used to preprocess the raw data
               'here', # this package can set the working directory as the root directory the script exist
               'papaja') # this package can create the manuscripts to APA.
```


set working directory and hyperparameter(subject index)

```{r}
# set directory
here::set_here() # set the working directory to the current root directory
idx <- formatC(4, width = 2, flag = "0") # translate the number as the format '0x'
folder_path <- paste('PreData/sub-', formatC(idx, width = 2, flag = "0"), sep = '') # create the root directory to storage preprocessed data
if(!dir.exists(folder_path)) {dir.create(folder_path)} # if not existing the directory, then create it.
```


# exp1

## preprocess for experiment 1

```{r}
# the parameter of the exp 1 
taskidx <- 1 # this is the number of task
filename <- paste('../2_Data/BIDS/sub-',idx,'/beh/','sub-',idx,'_task-GTDT_run-0',taskidx,'_beh.csv',sep = '') # the name of the file of experiment 1

# clean the raw data of exp1
df1 <- read.csv(filename)%>% # load the file 
      filter(!is.na(bg.started))%>% # remove the slide irrelevant of the experiment 
      mutate(Block=as.factor(blocks.thisTrialN), # the index of each block
             BlockIndex = as.factor(trials.thisTrialN), # the index in each block
             SubIndex = c(1:length(trials.thisTrialN))-1, # the index in each experiment
             CircleStartTime = as.numeric(annular.started), # the start time of circle
             CircleEndTime = as.numeric(annular.stopped), # the end time of circle
             SubjIdx = as.factor(idx), # the index of the participant
             FrameRate = as.factor(frameRate), # the frame rate of the screen 
             Version = as.factor(psychopyVersion), # the version of the psychopy
             Rt = as.factor(gsub("\\[|\\]", "", press.rt)), # capture the rt from circle onset
             SquareStartTime = as.numeric(ssec), # the start time of square
             SquareEndTime = as.numeric(esec), # the end time of square
             )%>%
      select(FrameRate,
             Version,
             SubjIdx,
             Block,
             BlockIndex,
             SubIndex,
             CircleStartTime,
             CircleEndTime,
             SquareStartTime,
             SquareEndTime,
             Rt
             )%>%
      separate(Rt, # separate the rts into different columns
               into = paste0("Rt", 1:2), 
               sep = ", ", 
               remove = T)%>%
      mutate(Rt1 = as.numeric(Rt1),
             Rt2 = as.numeric(Rt2),
             )%>%
      pivot_longer(cols = c(Rt1, Rt2), 
                   names_to = "RtTimes", 
                   names_pattern = "Rt?(.*)",
                   values_to = "Rt")%>% # separate the rts into different rows
      filter(!(is.na(Rt)&RtTimes==2))%>% # filter the trial with on response
  mutate(RtSquare = Rt - SquareStartTime) # calculate the rt from  square onset



# calculate the accuracy of of exp1 by each block 
df1.acc <- df1 %>%
  group_by(Block)%>%
  mutate(acc= ifelse(is.na(Rt),0,ifelse(Rt>1.6,0,1)))%>% # if rt is nan or >= 1.6, rt is 0
  summarise(acca=mean(acc))%>% # calculate the accuracy of each block
  ungroup()

# add a new column "acc" of each block
df1 <- merge(df1, df1.acc, by = "Block") # merge the accuracy and the data by block
write.csv(df1,paste('PreData/sub-',idx,'/sub-',idx,'_task-GTDT_run-0',taskidx,'_beh.csv',sep='')) # save the data
```


## the figure of exp1
```{r}
# the valid percentage of rt
# the valid percentage of rt is the first response of each trial
df1.tmp <- df1%>%
  filter(!RtTimes==2)
print(nrow(df1.tmp)/nrow(df1))
# examine the histogram of rt
ggplot(data=df1.tmp,mapping = aes(x = RtSquare))+
  geom_histogram(bins = 30)+
  theme_apa()
```



## mixed exp2 and exp3

note: exp2 and exp3 all contains the mixture of task 2 and task3, so we need to separate the task 2 and task 3 from experiment 2 and experiment 3.


```{r}
# load the raw data of experiment 1 
taskidx <- 2 # this is the number of experiment 2
filename <- paste('../2_Data/BIDS/sub-',idx,'/beh/','sub-',idx,'_task-GTDT_run-0',taskidx,'_beh.csv',sep = '')
exp1 <- read.csv(filename) # load the data

# load the raw data of experiment 2
taskidx <- 3 # this is the number of experiment 3
filename <- paste('../2_Data/BIDS/sub-',idx,'/beh/','sub-',idx,'_task-GTDT_run-0',taskidx,'_beh.csv',sep = '')
exp2 <- read.csv(filename) # load the data

# merge the two experiments
df <- dplyr::bind_rows(exp1,exp2) # merge the two experiments by rows

```



# exp2

## preprocess for experiment 2



```{r}
# this is the number of task
taskidx <- 2 

# select the data of task2
df2 <- df %>% # extract the experiment 2 from the merged data, experiment 3 of the column is NA
  filter(!is.na(key_resp_4.stopped))%>% # remove the slide irrelevant of the experiment 
  mutate(across(matches("^trials(?:_[0-9]+)?\\.thisTrialN$"),~replace(.,is.na(.),0)), # in this experiment, each block has its own columns, to get the index of the experiment, we need to translate the NA to 0, and accumulated them.
         BlockIndex = as.numeric(rowSums(across(matches("^trials(?:_[0-9]+)?\\.thisTrialN$")))), # block columns that don't belong to current block has the value 0, accumulated them is the index in each block.
         SubIndex = as.numeric(c(1:length(trials.thisTrialN))-1), # the index in each experiment
         across(matches("^annular.(?:_[0-9]+)?\\.started$"),~replace(.,is.na(.),0)), # in this experiment, each block has its own columns, to get the Circle start time of the experiment, we need to translate the NA to 0, and accumulated them.
         CircleStartTime = as.numeric(rowSums(across(matches("^annular.(?:_[0-9]+)?\\.started$")))), # annular columns that don't belong to current block has the value 0, accumulated them is the annular start in each block.
         across(matches("^annular.(?:_[0-9]+)?\\stopped$"),~replace(.,is.na(.),0)), # in this experiment, each block has its own columns, to get the Circle end time of the experiment, we need to translate the NA to 0, and accumulated them.
         CircleEndTime = as.numeric(rowSums(across(matches("^annular.(?:_[0-9]+)?\\stopped$")))), # annular columns that don't belong to current block has the value 0, accumulated them is the annular end in each block.
         SubjIdx = as.factor(idx), # the index of the participant
         FrameRate = as.factor(frameRate), # the frame rate of the screen 
         Version = as.factor(psychopyVersion), # the version of the psychopy
         Rt = as.numeric(gsub("\\[|\\]", "", key_resp_4.rt)), # capture the rt from circle onset
         Block = as.factor((as.integer(SubIndex - BlockIndex)/25)) # calculate the index of the block, each block has 25 trials.
         )%>%
      select(FrameRate,
             Version,
             SubjIdx,
             Block,
             BlockIndex,
             SubIndex,
             CircleStartTime,
             CircleEndTime,
             Rt
             )%>%
      separate(Rt, # separate the rts into different columns
               into = paste0("Rt", 1:2), 
               sep = ", ", 
               remove = T)%>%
      mutate(Rt1 = as.numeric(Rt1),
             Rt2 = as.numeric(Rt2),
             )%>%
      pivot_longer(cols = c(Rt1, Rt2), 
                   names_to = "RtTimes", 
                   names_pattern = "Rt?(.*)",
                   values_to = "Rt")%>% # separate the rts into different rows
      filter(!(is.na(Rt)&RtTimes==2)) # filter the trial with on response

# calculate the accuracy of experiment 2
df2.acc.2 <- df2 %>%
  mutate(acc= ifelse(is.na(Rt),0,ifelse(Rt>1.6,0,1)))%>% # if rt is nan or >= 1.6, rt is 0
  group_by(Block)%>%
  summarise(acca=mean(acc))%>% # calculate the accuracy of each block
  ungroup()
df2 <- merge(df2, df2.acc.2, by = "Block") # merge the accuracy and the data by block

# write the file
write.csv(df2,paste('PreData/sub-',idx,'/sub-',idx,'_task-GTDT_run-0',taskidx,'_beh.csv',sep='')) # save the data

```




## the figure of exp2
```{r}
# the valid percentage of rt
df2.tmp <- df2%>%
  filter(!RtTimes==2)
print(nrow(df2.tmp)/nrow(df2))
# the histogram of rt
ggplot(data=df2.tmp,mapping = aes(x = Rt))+
  geom_histogram(bins = 30)+
  theme_apa()
```


```{r}
# this is the number of task
taskidx <- 3 

# select the data of task3
df3 <- df %>% # extract the experiment 2 from the merged data, experiment 3 of the column is NA
  filter(is.na(key_resp_4.stopped)&!is.na(asec))%>% # remove the slide irrelevant of the experiment 
  mutate(across(matches("^trials(?:_[0-9]+)?\\.thisTrialN$"),~replace(.,is.na(.),0)),# in this experiment, each block has its own columns, to get the index of the experiment, we need to translate the NA to 0, and accumulated them.
         BlockIndex = as.numeric(rowSums(across(matches("^trials(?:_[0-9]+)?\\.thisTrialN$")))),# block columns that don't belong to current block has the value 0, accumulated them is the index in each block.
         SubIndex = as.numeric(c(1:length(trials.thisTrialN))-1), # the index in each experiment
         across(matches("^annular.(?:_[0-9]+)?\\.started$"),~replace(.,is.na(.),0)), # in this experiment, each block has its own columns, to get the Circle start time of the experiment, we need to translate the NA to 0, and accumulated them.
         CircleStartTime = as.numeric(rowSums(across(matches("^annular.(?:_[0-9]+)?\\.started$")))), # annular columns that don't belong to current block has the value 0, accumulated them is the annular start in each block.
         across(matches("^annular.(?:_[0-9]+)?\\stopped$"),~replace(.,is.na(.),0)), # in this experiment, each block has its own columns, to get the Circle end time of the experiment, we need to translate the NA to 0, and accumulated them.
         CircleEndTime = as.numeric(rowSums(across(matches("^annular.(?:_[0-9]+)?\\stopped$")))), # annular columns that don't belong to current block has the value 0, accumulated them is the annular end in each block.
         SubjIdx = as.factor(idx), # the index of the participant
         FrameRate = as.factor(frameRate), # the frame rate of the screen 
         Version = as.factor(psychopyVersion), # the version of the psychopy
         Block = as.factor(as.integer((SubIndex - BlockIndex)/25)) # calculate the index of the block, each block has 25 trials.
         )%>%
      select(FrameRate,
             Version,
             SubjIdx,
             Block,
             BlockIndex,
             SubIndex,
             CircleStartTime,
             CircleEndTime,
             )
# calculate the count 
ct <- df %>% # load the file 
      filter(!is.na(textbox.text))%>% # remove all slides without textbox
      mutate(Block=as.factor(c(1:length(trials.thisTrialN))-1))%>%
      select(Block,
             textbox.text)

# calculate the accuracy
bn<-df3%>%group_by(Block)%>%summarise(Count=max(BlockIndex))%>%ungroup()
df3.acc <- 1- sum(abs(ct$textbox.text - bn$Count))/sum(bn$Count)# calculate the accuracy, 1 - error rate, error rate is the deviation of actual count and set count
print(df3.acc)
# merge the data and accuracy
df3 <- merge(df3, ct, by = "Block")%>% # merge the accuracy and the data by block
      group_by(Block)%>%
      mutate(BlockIndex = as.numeric(BlockIndex),
             error = textbox.text - max(BlockIndex)-1)%>% # the deviation of actual count and set count
      ungroup()%>%
      arrange(SubIndex)

# write the file
write.csv(df3,paste('PreData/sub-',idx,'/sub-',idx,'_task-GTDT_run-0',taskidx,'_beh.csv',sep='')) # save the data
```



